Scaffolder - Software for \change{Reproducible}{Manual} Genome Scaffolding.

### 2011-08-15 src/article.tex
Capitalise Scaffolder consistently

The software outlined here, ``Scaffolder,'' aims to address these problems of
reproducibility by creating a file syntax and software framework for editing
a genome scaffold. Scaffolder uses a specific file format to define how
contigs are joined, additional sequences are inserted, and for the
specification of unknown regions. This syntax allows a scaffold to be updated
by simply editing the scaffold file. As such, scaffolder facilitates
a reproducible \change{scaffolding}{finishing} process and provides a concise
overview of how the final genomic scaffold was constructed. \pb

@@

Scaffolder is written in the Ruby programming language \change{using version
1.8.7}{and tested against versions 1.8.7 and 1.9.2} \cite{ruby-lang}. The
scaffolder package is split into two libraries. The first called ``scaffolder''
which provides the core Scaffolder application programming interface (API). The
second library ``scaffolder-tools'' provides the Scaffolder command line
interface (CLI). \pb

@@

The Scaffolder software facilitates reproducibly joining nucleotide sequences
together into a single contiguous scaffolded super-sequence. Plain-text
scaffold files written in \remove{the} YAML \remove{syntax} specify how these
sequences should be joined\change{. and the}{The} scaffolder software is used
to generate the scaffold sequence from these instructions. In addition to
specifying which contigs are required, the scaffold file allows the contigs to
be edited into smaller sub-sequences or reverse complemented if necessary.
\add{Each scaffold file represents one scaffolded nucleotide sequence and
separate scaffolds should be defined in separate files.} \pb

@@

The last entry in the scaffold, highlighted in blue, adds the sequence named
`sequence2' to the scaffold. This entry demonstrates how this sequence may be
manipulated prior to addition to the scaffold. The `start' and `stop' tags
trim the sequence to these coordinates inclusively.\remove{while} The
``reverse'' tag \add{also} instructs scaffolder to \remove{also} reverse
complement the sequence. In the putative scaffold shown in Figure 1 this
completes the \remove{scaffold } sequence. \pb

@@

Scaffolder provides a standardised set of Ruby classes and methods (API) for
interacting with the scaffold. This allows Scaffolder to be integrated into
existing genomics workflows or used with Ruby build tools such as Rake. In
addition Scaffolder provides a command line interface (CLI) to validate the
scaffold file and build the draft super sequence. The Scaffolder CLI behaves
as a standard Unix tool and returns appropriate exit codes and manual pages.
The use of both these Scaffolder interfaces is outlined in detail on the
scaffolder website (\scaffolder). \add{This website provides a ``getting
started'' guide as an introduction to using Scaffolder to build a genome
scaffold.} \pb

@@

\add{AGP is a similar format for describing scaffolds. This format can be used
to describe contig order and N-filled gap regions in a scaffold. The
advantages of the AGP format is that each contig entry is defined on a single
line which allows searching the scaffold using UNIX line-based tools. The
scaffolder format in comparison in contrast is written in the standardised
YAML format and therefore accessible to the many languages which provide
parsers to this format. The scaffolder format is provided with a tool
explicitly to produce the fasta sequence of scaffold specified by in the file.
The scaffold format further provides functions for trimming and replacing
regions of sequence using inserts.} \pb

@@

\add{An example use case for scaffolder is a combination of computational and
manual editing of a genome scaffold. We have used scaffolder in our own genome
projects to create an initial scaffold from compuationally parsing the output
of \emph{in silico} scaffolding tools into YAML. This scaffold was then
manually updated as the scaffold was finished with additional gap-filling
sequences generated in the laboratory.} \pb

>>>>

The software outlined here, ``Scaffolder,'' aims to address these problems of
reproducibility by creating a file syntax and software framework for editing
a genome scaffold. Scaffolder uses a specific file format to define how
contigs are joined, additional sequences are inserted, and for the
specification of unknown regions. This syntax allows a scaffold to be updated
by simply editing the scaffold file. As such, Scaffolder facilitates
a reproducible \change{scaffolding}{finishing} process and provides a concise
overview of how the final genomic scaffold was constructed. \pb

@@

Scaffolder is written in the Ruby programming language \change{using version
1.8.7}{and tested against versions 1.8.7 and 1.9.2} \cite{ruby-lang}. The
Scaffolder package is split into two libraries. The first called ``scaffolder''
which provides the core Scaffolder application programming interface (API). The
second library ``scaffolder-tools'' provides the Scaffolder command line
interface (CLI). \pb

@@

The Scaffolder software facilitates reproducibly joining nucleotide sequences
together into a single contiguous scaffolded super-sequence. Plain-text
scaffold files written in \remove{the} YAML \remove{syntax} specify how these
sequences should be joined\change{. and the}{The} Scaffolder software is used
to generate the scaffold sequence from these instructions. In addition to
specifying which contigs are required, the scaffold file allows the contigs to
be edited into smaller sub-sequences or reverse complemented if necessary.
\add{Each scaffold file represents one scaffolded nucleotide sequence and
separate scaffolds should be defined in separate files.} \pb

@@

The last entry in the scaffold, highlighted in blue, adds the sequence named
`sequence2' to the scaffold. This entry demonstrates how this sequence may be
manipulated prior to addition to the scaffold. The `start' and `stop' tags
trim the sequence to these coordinates inclusively.\remove{while} The
``reverse'' tag \add{also} instructs Scaffolder to \remove{also} reverse
complement the sequence. In the putative scaffold shown in Figure 1 this
completes the \remove{scaffold } sequence. \pb

@@

Scaffolder provides a standardised set of Ruby classes and methods (API) for
interacting with the scaffold. This allows Scaffolder to be integrated into
existing genomics workflows or used with Ruby build tools such as Rake. In
addition Scaffolder provides a command line interface (CLI) to validate the
scaffold file and build the draft super sequence. The Scaffolder CLI behaves
as a standard Unix tool and returns appropriate exit codes and manual pages.
The use of both these Scaffolder interfaces is outlined in detail on the
Scaffolder website (\scaffolder). \add{This website provides a ``getting
started'' guide as an introduction to using Scaffolder to build a genome
scaffold.} \pb

@@

\add{AGP is a similar format for describing scaffolds. This format can be used
to describe contig order and N-filled gap regions in a scaffold. The
advantages of the AGP format is that each contig entry is defined on a single
line which allows searching the scaffold using UNIX line-based tools. The
Scaffolder format in comparison in contrast is written in the standardised
YAML format and therefore accessible to the many languages which provide
parsers to this format. The Scaffolder format is provided with a tool
explicitly to produce the fasta sequence of scaffold specified by in the file.
The scaffold format further provides functions for trimming and replacing
regions of sequence using inserts.} \pb

@@

\add{An example use case for Scaffolder is a combination of computational and
manual editing of a genome scaffold. We have used Scaffolder in our own genome
projects to create an initial scaffold from compuationally parsing the output
of \emph{in silico} scaffolding tools into YAML. This scaffold was then
manually updated as the scaffold was finished with additional gap-filling
sequences generated in the laboratory.} \pb
